Optimizing Abstract Abstract Machines

### 2012-11-14 paper/paper.tex
Took a pass on abstracting compound literal data structures

Performance issues arise when one chooses the
``obvious,'' or ``natural'' abstraction for compound data structures like list. 
%
Our running
example will be the classic Boyer benchmark. The detail of note for
the Boyer benchmark is that it has large list literals, the exact
values of which do not matter for many analyses. If we
interpret a list literal of 30 elements (that again have 10 elements
each) as 300 applications of {\tt cons}, and then interpret a function
that traverses the tree, we will end up interpreting that function at
least 300 times, when once or twice would do.

We will first explain the natural abstraction of tuples, and then we
will explain a less precise allocation strategy from uniform $k$-CFA
that we use for large compound data literals that we modified from the
implementation in \citep{dvanhorn:wright-jagannathan-toplas98}.

\subsubsection{\Naively}

@@

\subsubsection{Less \naively{} and less precisely}

@@

For each benchmark, we collect analysis times, peak memory usage, and
the rate of states-per-second explored by the analysis for each of the
optimizations discussed in section~\ref{sec:opt}, culmulatively
applied.  The analysis is stopped after consuming 30 minutes of time
or 1 gigabyte of space.  When presenting \emph{relative} numbers, we
use the timeout limits as a lower bound on the actual time required,
thus giving a conservative estimate of improvements.

>>>>

The abstraction of compound data structures like lists has deep implications 
for precision and performance.
%
The classic list-heavy benchmark {\tt boyer}---a simplified implementation
of the Boyer-Moore theorem-prover---drove our own considerations for
abstractions for compound data structures.
%
Boyer's fluent use of literal list dooms the ``natural'' abstraction
of lists to fail from overprecision.
%
In short, if we interpret a literal list {\tt '(a b c)} as {\tt (cons 'a (cons 'b
(cons 'c '())))}, even a monovariant allocation stategy for abstract
cons cells will precisely and exactly create a 3-cell list.

If code contains a length-300 literal list, then its analysis yields a length-300 abstract
list as well.
%
In many cases, the specific contents of the list add little precision
to the resulting analysis, and yet the analyzer will dutifully execute
recursive functions over the entirety of these structures at every
encounter.
%
An alternative in this case is to explicitly flatten literal lists
into single abstract cells.

We will explore both options here.
%
We will first explain the natural abstraction of tuples, and then we
will explain a less precise allocation strategy from uniform $k$-CFA
that we use for large compound data literals that we modified from the
implementation in \citep{dvanhorn:wright-jagannathan-toplas98}.

\subsubsection{Option 1: The natural abstraction}

@@

\subsubsection{Option 2: A precise yet compact abstraction}

@@

For each benchmark, we collect analysis times, peak memory usage, and
the rate of states-per-second explored by the analysis for each of the
optimizations discussed in section~\ref{sec:opt}, cumulatively
applied.  The analysis is stopped after consuming 30 minutes of time
or 1 gigabyte of space.  When presenting \emph{relative} numbers, we
use the timeout limits as a lower bound on the actual time required,
thus giving a conservative estimate of improvements.
